---
title: "Fungal-clocks: Ophio v. Beau"
author: Biplabendu Das
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
    keep_md: no
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)

## For more inspiration on customizing the html output, refer to the following:
# https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents

```


```{r housekeeping, include=FALSE}
set.seed(420)
rm(list = ls())

## Load the libraries
pacman::p_load(pheatmap, dendextend, tidyverse, viridis)
pacman::p_load(RSQLite, tidyverse, dbplyr, DT, conflicted)
pacman::p_load(patchwork, glue)
pacman::p_load(circular)

## set conflict preference
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("layout", "plotly")
conflict_prefer("hclust", "flashClust")
conflict_prefer("union", "dplyr")

## set path to your working directory
path_to_repo = "/Users/biplabendudas/Documents/GitHub/Das_et_al_2022a"

## load functions
# customized theme for publication quality figures
source(paste0(path_to_repo,"/functions/theme_publication.R"))
# function to perform enrichment analysis
source(paste0(path_to_repo,"/functions/enrichment_analysis.R"))
# # function to plot z-scores (Cflo genes only)
source(paste0(path_to_repo,"/functions/plot_zscores.R"))

```


```{r set_parameters}

# SAMPLE NAME
## specify sample name
sample.name <- c("beau", "ophio_cflo", "ophio_kim")
## color scheme for the samples
col.scheme <- c("#5A829F", "#AD212F", "black")

# SCRIPT NAME
## specify the name of the script (folder) where figures will be saved
script.name <- "01_comparing_gene_exp_ophio_beau"

# eJTK OUTPUT
## Set GammaP threshold below which genes are classified as rhythmic
gamma.pval = 0.05
## Set false discovery rate for functional enrichment analyses
FDR = 5

```


```{r load_data}

# LOAD DATABASES (TC7)
# 1. TC6_ejtk.db
# Desc: This database contains all ejtk-output for TC6
ejtk.db <- dbConnect(RSQLite::SQLite(),
                     "./data/databases/TC6_fungal_ejtk.db")
# which tables are in the database
src_dbi(ejtk.db)
#
# 2. TC6_data.db
data.db <- dbConnect(RSQLite::SQLite(),
                     "./data/databases/TC6_fungal_data.db")
src_dbi(data.db)
#

```


## Overview/Goals

### 1. General patterns of gene expression

```{r general_exp}
# number of all genes
all.genes <- list()
for (i in 1:length(sample.name)) {
  all.genes[[i]] <- tbl(data.db, paste0(sample.name[[i]] ,"_fpkm")) %>%  
    collect()
  
  writeLines(paste("Number of genes in", sample.name[[i]], ":", nrow(all.genes[[i]])))
}

# A1: genes that have NO expression (FPKM == 0 at all time points)
not.expressed <- list()
for (i in 1:length(sample.name)) {
  not.expressed[[i]] <-
    tbl(data.db, paste0(sample.name[[i]] ,"_fpkm")) %>% 
    collect() %>% 
    filter_at(vars(starts_with("Z")), all_vars(. == 0)) %>%
    pull(gene_name)
  
  # How many genes are not expressed?
  writeLines(paste("n(genes-NOT-EXPRESSED) in", sample.name[[i]], ":", length(not.expressed[[i]])))
  
}

# A2: run enrichment (make plot of enrichment found of non-expressed genes)
for (i in 1:length(sample.name)) {
  writeLines(paste("running GO enrichment for NOT-EXPRESSED genes in", sample.name[[i]]))
  # run enrichment
  not.expressed[[i]] %>% 
    go_enrichment(., 
                  org = sample.name[[i]], 
                  bg = 'all') %>% # enrichment against all ophio_cflo genes in the genome
    
    # # pull gene names for a given GO term
    # separate_rows(., gene_name, sep = ", ") %>%
    # filter(GO == "GO:0009405") %>% # pathogenesis
    # # filter(GO == "GO:0090729") %>% # toxin activity
    # # filter(GO == "GO:0044419") %>% # interspecies interaction between organisms
    # # filter(GO == "GO:0020037") %>% # heme binding
    # pull()
    
    go_enrichment_plot(clean = "no") %>% 
    print()
  
}

# B: genes that are expressed (FPKM > 1 for at least one time point)
expressed <- list()
for (i in 1:length(sample.name)) {
  expressed[[i]] <- 
    tbl(data.db, paste0(sample.name[[i]],"_expressed_genes")) %>% 
    filter(expressed=="yes") %>% 
    collect() %>% 
    pull(gene_name) 
  
  # How many genes are expressed?
  writeLines(paste("n(EXPRESSED) in", sample.name[[i]], ":", length(expressed[[i]])))
}

```


### 2. Daily rhythms in gene expression

```{r daily_exp}

## Load all the rhythmic genesets 
## Note, ordered according to their p-value; highly rhythmic at the top.
#
# Choose period
period = '24'

## 
rhy <- list()
for (i in 1:2) {
  rhy[[i]] <-
    tbl(ejtk.db, paste0(sample.name[[i]],"_zscores_",period,'h')) %>%
    filter(GammaP < gamma.pval) %>%
    select(ID, GammaP) %>% collect() %>% arrange(GammaP) %>%
    select(ID) %>% pull()
  
  # How many genes are rythmic?
  writeLines(paste0("n(rhythmic-",period, "h) in ", sample.name[[i]], " : ", length(rhy[[i]])))
}

```


### Hierarchical clustering of rhythmic genes

- perform hierarchical clustering of 24h-rhythmic genes into four clusters;
- plot time-course heatmaps for the clustered 24h-rhythmic geneset
- Identify the day-peaking and night-peaking clusters visually.

```{r clustering_24h}

# Choose period
period = '24'

## 
rhy <- list()
for (i in 1:2) {
  rhy[[i]] <-
    tbl(ejtk.db, paste0(sample.name[[i]],"_zscores_",period,'h')) %>%
    filter(GammaP < gamma.pval) %>%
    select(ID, GammaP) %>% collect() %>% arrange(GammaP) %>%
    select(ID) %>% pull()
  
  # How many genes are rythmic?
  writeLines(paste0("n(rhythmic-",period, "h) in ", sample.name[[i]], " : ", length(rhy[[i]])))
}

## initialise lists to hold input and output of the hierarchical clustering
zscore.dat <- list() # zscore data (input)
my_gene_col <- list() # cluster identity for each rhythmic gene (output)
rhy.heat <- list() # pheatmap that can be saved/plotted (output)

# specify number of clusters
n_clusters <- 4

## run clustering and plot
for (i in 1:2) {
  ## load zscore dataset
  zscore.dat[[i]] <- data.db %>% tbl(., paste0(sample.name[[i]],"_zscores")) %>% collect()
  
  # Filter the zscores to keep only rhythmic genes
  zscore.rhy <-
    zscore.dat[[i]] %>% 
    filter(gene_name %in% rhy[[i]]) %>% 
    as.data.frame()
  
  # Set genes as rownames and convert it into a matrix
  rownames(zscore.rhy) = zscore.rhy$gene_name
  zscore.rhy <- as.matrix(zscore.rhy[-1])
  
  
  # Hierarchical clustering of the genesets
  my_hclust_gene <- hclust(dist(zscore.rhy), method = "complete")
  
  
  # Make annotations for the heatmaps
  my_clusters <- cutree(tree = as.dendrogram(my_hclust_gene), k = n_clusters) # k=  clusters
  my_gene_col[[i]] <- data.frame(cluster = my_clusters)
  
  
  # I’ll add some column annotations and create the heatmap.
  # Annotations for:
  # 1. Is the sample collected during the light or dark phase? 
  my_sample_col <- data.frame(phase = rep(c("light", "dark", "light"), c(5,6,1)))
  row.names(my_sample_col) <- colnames(zscore.rhy)
  
  # Manual color palette
  my_colour = list(
    phase = c(light = "#F2E205", dark = "#010440"),
    cluster = viridis::cividis(100)[c(10,90,60,30)]) #### NEED TO CHANGE #### account for n_clusters
  
  # Color scale
  my.breaks = seq(min(zscore.rhy), max(zscore.rhy), by=0.1)
  # my.breaks = seq(min(zscore.rhy), max(zscore.rhy), by=0.06)
  
  # Let's plot!
  rhy.heat[[i]] <-
    pheatmap(zscore.rhy, show_rownames = F, show_colnames = F,
             annotation_row = my_gene_col[[i]], 
             annotation_col = my_sample_col,
             cutree_rows = n_clusters, # OG was 4
             cutree_cols = 2,
             annotation_colors = my_colour,
             border_color=FALSE,
             cluster_cols = F,
             breaks = my.breaks,
             ## color scheme borrowed from: 
             color = inferno(length(my.breaks) - 1),
             # treeheight_row = 0, 
             # treeheight_col = 0,
             # remove the color scale or not
             main = paste0(sample.name[[i]], " 24h-rhythmic \n (n=", nrow(zscore.rhy), " genes)"),
             ## annotation legend
             annotation_legend = T,
             ## Color scale
             legend = T)
  
}

```


### Phase plots

```{r phase_plots}

rhy.24.sig <- list()
phase.ejtk <- list()

# Obtain the phases of 24h-rhythmic genes beau v. ophio_cflo
for (i in 1:2) {

rhy.24.sig[[i]] <- 
  tbl(ejtk.db, paste0(sample.name[i],"_zscores_24h")) %>% 
  filter(GammaP < gamma.pval) %>% 
  collect()

# Get the phases of the best matched waveforms
phase.ejtk[[i]] <- circular::circular(rhy.24.sig[[i]]$Phase, units="hours", template="clock24")
# # Get the time-of-day of expression peak
# phase.ejtk[[i]] <- circular::circular(rhy.24.sig[[i]]$MaxLoc, units="hours", template="clock24")
# # Get the time-of-day of expression trough
# phase.ejtk[[i]] <- circular::circular(rhy.24.sig[[i]]$MinLoc, units="hours", template="clock24")

}

# save all the circular phases in a list
l.phases <- phase.ejtk
# let's name the list elements for later use and reference
names(l.phases) <- sample.name[1:2]

writeLines("Performing Watson test to check if the average peak of 24h-rhythms in Beau and Ophio-cflo differs significantly")
# For all rhy genes
beau.ophio <- watson.two.test(l.phases[[1]],l.phases[[2]], alpha = FDR/100)
writeLines("Beau v. Ophio-cflo")
beau.ophio %>% print()

## Plot the phase distributions

# Initialize a list for saving the ggplots
g <- list()

means <- as.numeric(lapply(phase.ejtk, mean))
means <- circular(means, units="hours", template="clock24")


for(i in 1:length(l.phases)) {
  
  # define phase levels
  ordered_phases <- c("2","4","6","8","10","12",
                      "14","16","18","20","22","24")
    
  df.test <- l.phases[[i]] %>%
    as.data.frame() %>% 
    mutate(phase = x) %>%
    mutate(phase = replace(phase, x=="0", "24")) %>% 
    select(-x) %>% 
    group_by(phase = factor(phase, levels = ordered_phases)) %>%
    summarise(n_genes = n())

  m <- as.numeric(means[i])
    
  g[[i]] <-
    ggplot(df.test, aes(x=factor(phase), y=n_genes)) + 
    geom_bar(stat='identity', fill=col.scheme[[i]]) +
    
    xlab(c(names(l.phases)[i])) +
    
    scale_y_continuous(breaks = c(0,200,400,600)) +
    
    coord_polar() +
    theme_Publication() +
    theme(text = element_text(size = 15, colour = 'black'),
              # axis.title.x=element_blank(),
              # axis.text.x=element_blank(),
              legend.position = "none")
    #ggtitle(paste0("Dataset: ", names(l.phases)[i])) 
  
}

ggpubr::ggarrange(plotlist=g,
                  nrow = 2, ncol = 1,
                  widths = c(1,2), labels = NA)

```


### Clusters of rhythmic genes

- Which processes are identified clusters of rhythmic genes involved in?
- How do they fluctuate throughout the day?

#### Beau - 24h

```{r beau_clusters}
  
for (i in 1:n_clusters){
  
  writeLines(paste0("Species: ", sample.name[[1]], "\n", "24h-rhythmic genes, Cluster: ", i))
  
  # Summary
  genes <- my_gene_col[[1]] %>% rownames_to_column("g") %>% filter(cluster==as.character(i)) %>% pull(g)
  writeLines(paste0("n(genes) = ", length(genes),"\n"))
  
  # Enrichment
  overrepresented.terms <-
    genes %>% 
      go_enrichment(.,
                    function.dir = path_to_repo,
                    org = sample.name[[1]],
                    bg = expressed[[1]]) %>% 
      filter(adj_pVal < FDR/100) %>% 
      filter(over_under == "over")
  writeLines(paste0("\n", "n(overrepresented terms) = ", nrow(overrepresented.terms), "\n"))
  
  # Enriched terms word-cloud (borrowed from: https://towardsdatascience.com/create-a-word-cloud-with-r-bde3e7422e8a)
  if (nrow(overrepresented.terms)>0){
    # load libraries
    pacman::p_load(tm, wordcloud, RColorBrewer, wordcloud2)
    # get text as a character vector
    text <- overrepresented.terms %>% pull(GO_desc)
    # load your text data as a corpus
    docs <- Corpus(VectorSource(text)) # requires library "tm"
    # clean text (necessary?)
    docs <- docs %>%
              tm_map(removeNumbers) %>%
              tm_map(removePunctuation) %>%
              tm_map(stripWhitespace)
    docs <- tm_map(docs, content_transformer(tolower))
    docs <- tm_map(docs, removeWords, c("process", "molecular","cellular",
                                        "component", "compound", "part",
                                        "activity", "acid"
                                        ))
    # create document-term-matrix
    dtm <- TermDocumentMatrix(docs) 
    matrix <- as.matrix(dtm) 
    words <- sort(rowSums(matrix),decreasing=TRUE) 
    df <- data.frame(word = names(words),freq=words)
    # generate word-cloud
    wordcloud::wordcloud(words = df$word, freq = df$freq, min.freq = 2,
              max.words=200, random.order=FALSE, rot.per=0.35,
              scale=c(5,0.15),
              # colors=brewer.pal(8, "Dark2")
              colors=col.scheme[[1]]
              )
  }
  # Stacked zplot
  genes %>% 
  stacked.zplot_tc6(cond = "beau") %>% 
    multi.plot(rows = 1, cols = 1)
  
}

```

#### Ophio-cflo - 24h

```{r ophio_clusters}
  
for (i in 1:n_clusters){
  
  writeLines(paste0("Species: ", sample.name[[2]], "\n", "24h-rhythmic genes, Cluster: ", i))
  
  # Summary
  genes <- my_gene_col[[2]] %>% rownames_to_column("g") %>% filter(cluster==as.character(i)) %>% pull(g)
  writeLines(paste0("n(genes) = ", length(genes),"\n"))
  
  # Enrichment
  overrepresented.terms <-
    genes %>% 
      go_enrichment(.,
                    function.dir = path_to_repo,
                    org = sample.name[[2]],
                    bg = expressed[[2]]) %>% 
      filter(adj_pVal < FDR/100) %>% 
      filter(over_under == "over")
  writeLines(paste0("\n", "n(overrepresented terms) = ", nrow(overrepresented.terms), "\n"))
  
  # Enriched terms word-cloud (borrowed from: https://towardsdatascience.com/create-a-word-cloud-with-r-bde3e7422e8a)
    if (nrow(overrepresented.terms) > 0) {
      # load libraries
      pacman::p_load(tm, wordcloud, RColorBrewer, wordcloud2)
      # get text as a character vector
      text <- overrepresented.terms %>% pull(GO_desc)
      # load your text data as a corpus
      docs <- Corpus(VectorSource(text)) # requires library "tm"
      # clean text (necessary?)
      docs <- docs %>%
                tm_map(removeNumbers) %>%
                tm_map(removePunctuation) %>%
                tm_map(stripWhitespace)
      docs <- tm_map(docs, content_transformer(tolower))
      docs <- tm_map(docs, removeWords, c("process", "molecular","cellular",
                                          "component", "compound", "part",
                                          "activity", "acid"
                                          ))
      # create document-term-matrix
      dtm <- TermDocumentMatrix(docs) 
      matrix <- as.matrix(dtm) 
      words <- sort(rowSums(matrix),decreasing=TRUE) 
      df <- data.frame(word = names(words),freq=words)
      # generate word-cloud
      wordcloud::wordcloud(words = df$word, freq = df$freq, min.freq = 2,
                max.words=200, random.order=FALSE, rot.per=0.35,
                scale=c(4,0.15),
                # colors=brewer.pal(8, "Dark2")
                colors=col.scheme[[2]]
                )
    }
  
  
  # Stacked zplot
  genes %>% 
  stacked.zplot_tc6(cond = "ophio_cflo") %>% 
    multi.plot(rows = 1, cols = 1)
  
}

```