---
title: "compile_gene_exp_by_sample"
author: "Biplabendu Das"
date: "8/13/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Species: Ophiocordyceps camponoti-floridani

00_Housekeeping

- Read all the normalized gene expression files for Ophiocordyceps camponoti-floridani (ophio_cflo)

```{r}
# Absolute path to the data
path_to_data = "/Users/biplabendudas/Documents/GitHub/Das_et_al_2022a/results/normalized_gene_exp/ophio_cflo/"

# load libraries
library(data.table)
library(tidyverse)

# list all the normalized_gene_exp files (n = 12; Ophio_cflo)
data_all=list.files(path = path_to_data,
                    pattern='*.txt')
```

01_Read the files, extract the columns of interest, and compile them into one table

01.1
Extract two columns 
(gene_name "gene" and sample_name_gene_exp "value_2")

```{r}

foo <- data.frame()

for (i in 1:length(data_all)) {
  
  if (i == 1) {
  foo = read.csv(file = paste0(path_to_data,data_all[i]), header = T, sep = "\t") %>%
          select(gene_name=gene, value_2)

  names(foo)[2] <- paste0("ZT",
                          str_split(str_split(data_all[i], "_")[[1]][5], ".txt")[[1]][1])
  }
  
  else {
    
    bar = read.csv(file = paste0(path_to_data,data_all[i]), header = T, sep = "\t") %>%
          select(gene_name=gene, value_2)
    
    names(bar)[2] <- paste0("ZT",
                            str_split(str_split(data_all[i], "_")[[1]][5], ".txt")[[1]][1])
    
    foo <- foo %>% 
            left_join(bar, by="gene_name")
    
  }

}

# Take a look at the compiled table

foo %>% head()

# Save the compiled dataframe "foo"
write.csv(foo,
          file = paste0(path_to_data,"normalized_gene_exp_ophio_cflo_all_samples.csv"),
          row.names = F)

```


# Species: Ophiocordyceps kimflemingae
  # experiment: light-dark (ld)

00_Housekeeping

- Read all the normalized gene expression files for Ophiocordyceps kimflemingae (ophio_kim)

```{r}
# Absolute path to the data
path_to_data = "/Users/biplabendudas/Documents/GitHub/Das_et_al_2022a/results/normalized_gene_exp/ophio_kim/ld/"

# load libraries
library(data.table)
library(tidyverse)

# list all the normalized_gene_exp files (n = 12; Ophio_cflo)
data_all=list.files(path = path_to_data,
                    pattern='*.txt')
```

01_Read the files, extract the columns of interest, and compile them into one table

01.1
Extract two columns 
(gene_name "gene" and sample_name_gene_exp "value_2")

```{r}

foo <- data.frame()

for (i in 1:length(data_all)) {
  
  if (i == 1) {
  foo = read.csv(file = paste0(path_to_data,data_all[i]), header = T, sep = "\t") %>%
          select(gene_name=gene, value_2)

  names(foo)[2] <- paste0("ZT",
                          # extract the second sample's timepoint
                          str_split(str_split(data_all[i], "_")[[1]][7], ".txt")[[1]][1],
                          # extract the experiment's light regime
                          str_split(data_all[i], "_")[[1]][6])
  }
  
  else {
    
    bar = read.csv(file = paste0(path_to_data,data_all[i]), header = T, sep = "\t") %>%
          select(gene_name=gene, value_2)
    
    names(bar)[2] <- paste0("ZT",
                            str_split(str_split(data_all[i], "_")[[1]][7], ".txt")[[1]][1],
                            str_split(data_all[i], "_")[[1]][6])
    
    foo <- foo %>% 
            left_join(bar, by="gene_name")
    
  }

}

# Take a look at the compiled table

foo %>% head()

# Save the compiled dataframe "foo"
write.csv(foo,
          file = paste0(path_to_data,"normalized_gene_exp_ophio_kim_ld_samples.csv"),
          row.names = F)

```


# Species: Ophiocordyceps kimflemingae
  # experiment: constant dark or dark-dark (dd)

00_Housekeeping

- Read all the normalized gene expression files for Ophiocordyceps kimflemingae (ophio_kim)

```{r}
# Absolute path to the data
path_to_data = "/Users/biplabendudas/Documents/GitHub/Das_et_al_2022a/results/normalized_gene_exp/ophio_kim/dd/"

# load libraries
library(data.table)
library(tidyverse)

# list all the normalized_gene_exp files (n = 12; Ophio_cflo)
data_all=list.files(path = path_to_data,
                    pattern='*.txt')

data_all
```

01_Read the files, extract the columns of interest, and compile them into one table

01.1
Extract two columns 
(gene_name "gene" and sample_name_gene_exp "value_2")

```{r}

foo <- data.frame()

for (i in 1:length(data_all)) {
  
  if (i == 1) {
  foo = read.csv(file = paste0(path_to_data,data_all[i]), header = T, sep = "\t") %>%
          select(gene_name=gene, value_2)

  names(foo)[2] <- paste0("ZT",
                          # extract the second sample's timepoint
                          str_split(str_split(data_all[i], "_")[[1]][7], ".txt")[[1]][1],
                          # extract the experiment's light regime
                          str_split(data_all[i], "_")[[1]][6])
  }
  
  else {
    
    bar = read.csv(file = paste0(path_to_data,data_all[i]), header = T, sep = "\t") %>%
          select(gene_name=gene, value_2)
    
    names(bar)[2] <- paste0("ZT",
                            str_split(str_split(data_all[i], "_")[[1]][7], ".txt")[[1]][1],
                            str_split(data_all[i], "_")[[1]][6])
    
    foo <- foo %>% 
            left_join(bar, by="gene_name")
    
  }

}

# Take a look at the compiled table

foo %>% head()

# Save the compiled dataframe "foo"
write.csv(foo,
          file = paste0(path_to_data,"normalized_gene_exp_ophio_kim_dd_samples.csv"),
          row.names = F)

```

TO DO:
- All Beauveria bassiana samples (ld only)